---
title: "Lab5_Demo2_key"
author: "Mateo Robbins"
date: "2023-02-15"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(vip) #variable importance
```
## R
```{r}
#Grab the .csv from here:
#https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify
kaggle_dat <- read_csv("kaggle_genres_v2.csv")
unique(kaggle_dat$genre)
table(kaggle_dat$genre)

#Removing inappropriate columns and selecting trap and Hiphop as the two genres here and making case consistent

genre_dat <- kaggle_dat %>%
  select(-c(type, uri, track_href, analysis_url, `Unnamed: 0`, title, tempo, id, song_name)) %>%
  filter(genre == "Hiphop"|genre == "Rap") %>%
  mutate(genre = str_replace(genre, "Hiphop", "hiphop")) %>%
  mutate(genre = str_replace(genre, "Rap", "rap")) %>%
  mutate(genre = as.factor(genre))
```

```{r}
##split the data
genre_split <- initial_split(genre_dat)
genre_train <- training(genre_split)
genre_test <- testing(genre_split)
```

```{r recipe}
#Preprocess the data
genre_rec <- recipe(genre ~., data = genre_train) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_normalize(all_numeric(), -all_outcomes(),)
  
```

Set up a decision tree specification. Note: the cost_complexity parameter is a pruning penalty parameter that controls how much we penalize the number of terminal nodes in the tree.  It's conceptually similar to lambda from regularized regression.

```{r tree_specification}
tree_spec_fixed <- decision_tree(
  cost_complexity = 0.1,
  tree_depth = 4,
  min_n = 11
) %>%
  set_engine("rpart") %>%
  set_mode("classification")
```

But, as usual, we don't want just any old values for our hyperparameters, we want optimal values.
```{r}
#new spec, tell the model that we are tuning hyperparams
tree_spec_tune <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()) %>%
  set_engine("rpart") %>%
  set_mode("classification")

tree_spec_tune

tree_grid <- grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)
tree_grid
```

```{r workflow_tree}
wf_tree_tune <- workflow() %>%
  add_recipe(genre_rec) %>%
    add_model(tree_spec_tune)
```

```{r resampling}
#set up k-fold cv. This can be used for all the algorithms
genre_cv = genre_train %>% vfold_cv(v = 10)
genre_cv
```

```{r}
doParallel::registerDoParallel() #build trees in parallel
#As models become more complex, it is useful to measure running time
system.time(
tree_rs <- tune_grid(
  tree_spec_tune,
  genre ~., 
  resamples = genre_cv,
    grid = tree_grid,
  metrics = metric_set(accuracy)
)
)
tree_rs
```
Use autoplot() to examine how different parameter configurations relate to accuracy 
```{r}
autoplot(tree_rs) + theme_light()
```

```{r select_hyperparam}
show_best(tree_rs)
select_best(tree_rs)
```

We can finalize the model specification where we have replaced the tune functions with optimized values.

```{r final_tree_spec}
final_tree <- finalize_model(tree_spec_tune, select_best(tree_rs))
```

This model has not been fit yet though.

```{r final_tree_fit}
#similar functions here.
final_tree_fit <- fit(final_tree, genre~.,genre_train)
#last_fit() fits on training data (like fit()), but then also evaluates on the test data.
final_tree_result <- last_fit(final_tree, genre~., genre_split )
final_tree_result$.predictions 
```

#Visualize variable importance.
```{r tree_vip}
final_tree_fit %>%
  vip(geom = "col", aesthetics = list(fill = "midnightblue", alpha = 0.8)) +
  scale_y_continuous(expand = c(0,0))

```


